Type Error Telemetry at Scale

### 2023-09-30 programming23/paper.tex
shrinking 6

@@

Type error telemetry in Roblox Studio has two distinguishing characteristics:
it includes no personal information (PII), and it sends data on
randomly-selected runs of the type checker rather than specific events of interest.
To the best of our knowledge, this design occupies a unique position relative to prior
work~(\cref{t:telemetry-design}).

In other professional-grade IDEs, such as VSCode~\cite{vscode-telemetry},
IntelliJ~\cite{intellij-telemetry}, and \code{.NET}~\cite{dotnet-telemetry}, telemetry may include all sorts of
data: error reports, source code fragments, timestamps, and filepaths.
Users may be able to opt out of some telemetry, but the details depend on the
license agreement.
Furthermore, at least in VSCode, IDE extensions can report their own telemetry.
While this data can be invaluable for discovering bugs in production, it
must be handled with extreme care.

The Blackbox project similarly collects telemetry in broad strokes.
Every Java program submitted for compilation goes into the Blackbox database (minus
any comments above the main class) along with the compiler's response~\cite{bkmu-sigcse-2014}.
%% Though somewhat intrusive,
The millions of programs in the Blackbox dataset
have enabled dozens of contributions to the area of CS
education~\cite{bask-icer-2018}, ranging from categorization of novice
errors~\cite{mk-fie-2014,m-masters-2016} to analyses of error frequency and
time-to-fix~\cite{ab-sigcse-2015}.
Related prior work used the GRUMPS telemetry system to collect 4.7 million
keystroke-level actions from learners programming in
Ada~\cite{tm-iticse-2004,tkdmceg-ascilite-2003}.
This data is not available.

At the opposite end of the privacy spectrum, Go's design for \emph{transparent
telemetry} reports only counter values~\cite{transparent-telemetry}.
Unlike Roblox Studio, transparent telemetry includes no timestamps and no 
session ids.
The data is useful for learning about the frequency of specific events and the
distribution of events across clients (e.g., how many \code{CodeTooComplex}
errors arise? what is the average number of type errors?).
However, the lack of timestamps and ids means that it cannot track edits
over time (a feature that is quite useful to us, for example, in~\cref{f:error-density}).
% In~\cref{f:error-density}, for example, we see that increases to the number
% of type errors are quickly followed by decreases.
% Without timestamps, we would be unable to rule out the possibility that users
% build up a mountain of type errors and then slowly reduce them to zero.

>>>>

% TODO BEFORE SUBMISSION
% be consistent between "IDs" and "ids" [SK prefers "IDs" to avoid it being read as a word]

% TODO BEFORE FINAL
% - uncomment acks

@@

Our telemetry has two distinguishing characteristics:
it includes no private information (both PII and source code details),
and it sends data on
randomly-selected runs of the type checker rather than specific events of interest.
To the best of our knowledge, this design occupies a unique position relative to prior
work~(\cref{t:telemetry-design}).\todo{Where do these labels come
  from?!? There are no ``prior work'' citations either!}

In other professional-grade IDEs, such as VSCode~\cite{vscode-telemetry},
IntelliJ~\cite{intellij-telemetry}, and \code{.NET}~\cite{dotnet-telemetry}, telemetry may include all sorts of
data: error reports, source code fragments, time\-stamps, and filepaths.
Users may be able to opt out of some telemetry, but the details depend on the
license agreement.
Furthermore, at least in VSCode, IDE extensions can report their own telemetry.
While this data can be invaluable for discovering bugs in production, it
must be handled with extreme care.

The Blackbox project similarly collects telemetry in broad strokes.
Every Java program submitted for compilation goes into the Blackbox database (minus
any comments above the main class) along with the compiler's response~\cite{bkmu-sigcse-2014}.
%% Though somewhat intrusive,
The millions of programs in the Blackbox dataset
have enabled dozens of contributions to the area of CS
education~\cite{bask-icer-2018}. However, these operate at the
opposite end of privacy.

% BEN: saving space.
% , ranging from categorization of novice
% errors~\cite{mk-fie-2014,m-masters-2016} to analyses of error frequency and
% time-to-fix~\cite{ab-sigcse-2015}.
% Related prior work used the GRUMPS telemetry system to collect 4.7 million
% keystroke-level actions from learners programming in
% Ada~\cite{tm-iticse-2004,tkdmceg-ascilite-2003}.
% This data is not available.
% BEN: Not sure what the above sentence means and whether we should
% keep it. Is it that you can't freely get Blackbox, you have to ask
% Neil Brown for permission or something?

At the opposite end of the privacy spectrum, Go's design for \emph{transparent
telemetry} reports only counter values~\cite{transparent-telemetry}.
Unlike Roblox Studio, transparent telemetry includes no timestamps and no 
session ids.
While useful for learning about the frequency and distribution of
specific events, the lack of timestamps and ids means it cannot track
edits over time (which we use in \cref{f:error-density}).
% BEN: commented for space.
% it cannot track 
% The data is useful for learning about the frequency of specific events and the
% distribution of events across clients (e.g., how many \code{CodeTooComplex}
% errors arise? what is the average number of type errors?).
% However, the lack of timestamps and ids means that it cannot track edits
% over time (a feature that is quite useful to us, for example, in~\cref{f:error-density}).
% In~\cref{f:error-density}, for example, we see that increases to the number
% of type errors are quickly followed by decreases.
% Without timestamps, we would be unable to rule out the possibility that users
% build up a mountain of type errors and then slowly reduce them to zero.
